---
title: "FI vs FP imputation - HRS"
output: html_notebook
---
Glen Pridham Apr 2024

Script for imputing HRS data to use in FI vs FP

MICE. Algorithm(s): 
-causal CART
-maybe 2-level model - too slow
  -try logistic regression
-maybe that's enough?

```{r}
outputDir = "/home/glen/analysis/fi_vs_fp"
gRootDir = "/home/glen/Documents/r" #where scripts are
gOutputDir=outputDir

setwd(outputDir)
print(getwd())

source(sprintf("%s/nhanes.R",gRootDir),verbose=0)
source(sprintf("%s/goodness_of_fit.R",gRootDir),verbose=0)
source(sprintf("%s/pca.R",gRootDir),verbose=0)
source(sprintf("%s/pca_fi.R",gRootDir),verbose=0)
library("mice")
```



```{r}
codex = read.csv(sprintf("%s/rand_fi.csv",outputDir),row.names=1)
```

```{r}
fiVar = c("nhmliv", "shlt",   "hibpe",  "diabe",  "cancre", "lunge",  "hearte", "stroke", "arthre",
          "walksa", "walk1a", "chaira", "climsa", "clim1a", "stoopa", "armsa",  "pusha",  "lifta",
          "dimea","dressa", "dressh", "walkra", "walkre", "walkrh", "batha",  "bathh",  "eata",
          "eath",   "beda",   "bede",   "bedh",   "toilta", "toilth", "shopa",  "phonea", "moneya",
          "hosp",   "nrshom", "outpt",  "homcar", "spcfac")
names(fiVar) = codex[fiVar,"Label"]

fpVar = c('fp1_weight','fp2_grip','fp3_gait','fp4_exhaustion','fp5_low_activity')

auxVar = c("age","sex") #"walk_time_ave","height","weight" # I'm worried about overfitting #"fi_prev" probably good but I'm concerned I'll get pushback
```



```{r}
#8 is first wave, but keep 7 for carry forward / imputation
  #nevermind, wave 7 has no FP data in it
waves = 8:14 #no covid (I forget what's missing in early, some measurement like grip or gait)
keepWaves = rep(T,length(waves))
#keepWaves[1] = F #drop since no prev?
keepWaves[1:2+length(keepWaves)-2] = F #won't have 'next'
years = 1992+2*(waves-1)
```

FI analysis - df and d2f


```{r}
long = list()
for (i in 1:length(waves))
{

  long[[i]] = read.csv(sprintf("%s/data/hrs_rand_preproc_wave%02d.csv",gOutputDir,waves[i]),row.names=1)
  #must have age and FI
  
  #long[[i]][,sprintf("%s_next",fiVar)] = NA
  #long[[i]][,sprintf("%s_next",fpVar)] = NA
  #long[[i]][,"fi_next"] = NA
  #long[[i]][,"Nfp_next"] = NA
  long[[i]][,sprintf("%s_prev",fiVar)] = NA
  long[[i]][,sprintf("%s_prev",fpVar)] = NA
  long[[i]][,"fi_prev"] = NA
  long[[i]][,"Nfp_prev"] = NA
  long[[i]] = subset(long[[i]],!is.na(fi) & !is.na(age))
  if(i> 1)
  {
    ids = intersect(rownames(long[[i-1]]),rownames(long[[i]]))
    #long[[i-1]][ids,sprintf("%s_next",fiVar)] = long[[i]][ids,fiVar]
    #long[[i-1]][ids,"fi_next"] = apply(long[[i]][ids,fiVar],1,mean,na.rm=T)
    #long[[i-1]][ids,sprintf("%s_next",fpVar)] = long[[i]][ids,fpVar]
    #long[[i-1]][ids,"fp_next"] = apply(long[[i]][ids,fpVar],1,sum,na.rm=T)
    
    long[[i]][ids,sprintf("%s_prev",fiVar)] = long[[i-1]][ids,fiVar]
    long[[i]][ids,"fi_prev"] =          apply(long[[i-1]][ids,fiVar],1,mean,na.rm=T)
    availfpVar = intersect(colnames(long[[i-1]]),fpVar)
    print(availfpVar)
    long[[i]][ids,sprintf("%s_prev",availfpVar)] = long[[i-1]][ids,availfpVar]
    long[[i]][ids,"Nfp_prev"] =          apply(long[[i-1]][ids,availfpVar],1,mean,na.rm=T)*length(availfpVar)

  }
  
}
for (i in 1:length(long))
{
  if(i <= (length(long)-2)) #look forward by 2 since gait/grip are never measured back-to-back
  {
    ids = intersect(rownames(long[[i+2]]),rownames(long[[i]]))
    long[[i]][ids,sprintf("%s_next",fiVar)] = long[[i+2]][ids,fiVar]
    long[[i]][ids,"fi_next"] =          apply(long[[i+2]][ids,fiVar],1,mean,na.rm=T)
    availfpVar = intersect(colnames(long[[i+2]]),fpVar)
    print(availfpVar)
    long[[i]][ids,sprintf("%s_next",availfpVar)] = long[[i+2]][ids,availfpVar]
    long[[i]][ids,"Nfp_next"] =          apply(long[[i+2]][ids,availfpVar],1,mean,na.rm=T)*length(availfpVar)

  }
}
long = do.call(rbind,long[keepWaves])

#convert months to years
long[,"age"] = long[,"age"]/12
long[,"age_prev"] = long[,"age_prev"]/12
long[,"age_prev_prev"] = long[,"age_prev_prev"]/12
long[,"age_next"] = long[,"age_next"]/12
long[,"death_age"] = long[,"death_age"]/12

#convert sex to char
sex = rep("unknown",nrow(long))
sex[long[,"ragender"]==1] = "male"
sex[long[,"ragender"]==2] = "female"
long[,"sex"] = sex

logi = is.na(long[,"fp2_grip"]) & is.na(long[,"fp3_gait"])
print(sprintf("dropping %d entries with no grip and no gait",sum(logi))) #must have at least one
long = long[!logi,]
```


```{r}
hist(long[,"age"])
```
```{r}
long[,"id"] = long[,"rahhidpn"]
long[,"id"] = as.integer(long[,"id"]) #mice wants integer here or it gets mad

long = long[,c("id","rahhidpn","wave","age","sex",fiVar,fpVar,sprintf("%s_prev",fiVar),sprintf("%s_prev",fpVar))]

#'fix' types
for (j in 1:length(fiVar))
{
  long[,fiVar[j]] = as.integer(long[,fiVar[j]])
  long[,sprintf("%s_prev",fiVar[j])] = as.integer(long[,sprintf("%s_prev",fiVar[j])])
  #long[,fiVar[j]] = ordered(long[,fiVar[j]])
  #long[,sprintf("%s_prev",fiVar[j])] = ordered(long[,sprintf("%s_prev",fiVar[j])])
}
for (j in 1:length(fpVar))
{
  long[,fpVar[j]] = as.integer(long[,fpVar[j]])
  long[,sprintf("%s_prev",fpVar[j])] = as.integer(long[,sprintf("%s_prev",fpVar[j])])
  #long[,fpVar[j]] = ordered(long[,fpVar[j]])
  #long[,sprintf("%s_prev",fpVar[j])] = ordered(long[,sprintf("%s_prev",fpVar[j])])
}
```

```{r}

contImp = NULL
binImp = setdiff(c(fiVar,fpVar,sprintf("%s_prev",c(fiVar,fpVar))),contImp)

pred = make.predictorMatrix(long)
pred[,"wave"] = 0 #don't predict wave


#basic linear models
meth = make.method(long)
meth[1:length(meth)] = ""
meth[binImp] = "logreg"
meth[contImp] = "norm" #"2l.norm" #"2l.norm" #2l.pan doesn't allow missing values in predictors
print(meth)
```

```{r}
predCart = make.predictorMatrix(long)

methCart = make.method(long,defaultMethod="cart")
methCart["wave"] = ""
methCart[c(binImp,contImp)] = "cart"

print(methCart) #some may have no NAs in which case they default to ""
```

split linear
```{r}
save = T
load = T
file = sprintf("%s/data/hrs_rand_fi_vs_fp_mice_lin_fp.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_fp = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),"age","sex","wave","id")
  ml_fp = mice(long[,vars], meth = meth[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_fp,file)
}


file = sprintf("%s/data/hrs_rand_fi_vs_fp_mice_lin_fi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_fi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id")
  ml_fi = mice(long[,vars], meth = meth[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_fi,file)
}


file = sprintf("%s/data/hrs_rand_fi_vs_fp_mice_lin_fpfi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_fpfi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id")
  ml_fpfi = mice(long[,vars], meth = meth[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_fpfi,file)
}

```

check
what can we test?
-individual trajectories
-binary outcome frequencies
-Nfp / FI trajectories

start with individual variables
```{r}
library(colorspace)
fp = list()
C = list() #correlation
Cv = list() #covariance
Cvi = list() #covariance - imputed values only
mu = list()
age_cuts = seq(min(long[,"age"])-1,max(long[,"age"])+1,length=10)
agg = list()
aggse = list()
agg_fpvar = list() #one per fpvar
agg_fpvar_se = list()
for (j in 1:length(fpVar)) 
{
  agg_fpvar[[j]] = list()
  agg_fpvar_se[[j]]  = list()
}
for (i in 1:ml_fp$m)
{
  fp[[i]] = mice::complete(ml_fp,i)[,c("id","age","sex","wave",fpVar)]
  fp[[i]][,"imp"] = i
  fp[[i]][,"Nfp"] = apply(fp[[i]][,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,"Nfp_noimp"] = apply(long[,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,sprintf("%s_imputed",fpVar)] = ml_fp$where[,fpVar]
  fp[[i]][,"N_imputed"] = apply(ml_fp$where[,fpVar],1,sum)
  C[[i]] = cor(mice::complete(ml_fp,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))])
  Cv[[i]] = cov(mice::complete(ml_fp,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))])
  temp = mice::complete(ml_fp,i)[,c(fpVar,sprintf("%s_prev",fpVar))]
  temp[!ml_fp$where[,c(fpVar,sprintf("%s_prev",fpVar))]] = NA
  Cvi[[i]] = cov(temp,use='pairwise.complete')
  mu[[i]] = apply(mice::complete(ml_fp,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))],2,mean,na.rm=T)
  m = matrix(NA,nrow=length(mu[[i]]),ncol=2)
  m[,1] = mu[[i]]
  m[,2] = mu[[i]]
  mu[[i]] = m
  
  
  fp[[i]][,"age_cut"] = as.numeric(as.character(cut(fp[[i]][,"age"],age_cuts,labels=age_cuts[-1]/2+age_cuts[-length(age_cuts)]/2,include.lowest=T)))
  
  vars = c("age",fpVar,"Nfp","Nfp_noimp")
  agg[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
  aggse[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
  agg[[i]][,"N"] = aggregate(fp[[i]][,"age"],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
  agg[[i]][,"sex"] = as.numeric(factor(agg[[i]][,"sex"]),c("male","female"))-1
  aggse[[i]][,"N"] = 0
  aggse[[i]][,"sex"] = 0

  for(j in 1:length(fpVar))
  {
    agg_fpvar[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml_fp$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
    agg_fpvar_se[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml_fp$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
    agg_fpvar[[j]][[i]][,"N"] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml_fp$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
    agg_fpvar[[j]][[i]][,"sex"] = as.numeric(factor(agg_fpvar[[j]][[i]][,"sex"]),c("male","female"))-1
    agg_fpvar_se[[j]][[i]][,"N"] = 0
    agg_fpvar_se[[j]][[i]][,"sex"] = 0
  }

}
fp = do.call(rbind,fp)
C = RubinMat(C)
Cv = RubinMat(Cv)
Cvi = RubinMat(Cvi)
mu = RubinMat(mu)

agg = RubinMat(agg,lse=aggse)
vars = c(fpVar,"Nfp","Nfp_noimp")
agg[[1]] = as.data.frame(agg[[1]])
agg[[1]][,sprintf("%s_se",vars)] = agg[[2]][,vars]
agg = agg[[1]]

for (j in 1:length(fpVar))
{
  agg_fpvar[[j]] = RubinMat(agg_fpvar[[j]],agg_fpvar_se[[j]])
  agg_fpvar[[j]][[1]] = as.data.frame(agg_fpvar[[j]][[1]])
  agg_fpvar[[j]][[1]][,sprintf("%s_se","x")] = agg_fpvar[[j]][[2]][,"x"]
  agg_fpvar[[j]] = agg_fpvar[[j]][[1]]
}



#
ggplot(subset(agg,sex==0),aes(x=age_cut,y=fp1_weight,ymin=fp1_weight-fp1_weight_se,ymax=fp1_weight+fp1_weight_se,color=as.ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F,method="lm")+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red") #Blue-Red 3

ggplot(subset(agg,sex==1),aes(x=age,y=fp1_weight,ymin=fp1_weight-fp1_weight_se,ymax=fp1_weight+fp1_weight_se,color=as.ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F,method="lm")+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red") #Blue-Red 3

```

```{r}
vars = fpVar
xlim = c(0,.5)
g = list()
g[[1]] = TilePlot(cov(long[,vars],use='complete.obs'),range=xlim)+ggtitle("Complete-case")
g[[2]] = TilePlot(cov(long[,vars],use='pairwise.complete'),range=xlim)+ggtitle("Available-case")
g[[3]] = TilePlot(Cv[[1]][vars,vars],dm=Cv[[2]][vars,vars],range=xlim)+ggtitle("Imputed - logreg")
g[[4]] = TilePlot(Cvi[[1]][vars,vars],dm=Cvi[[2]][vars,vars],range=xlim)+ggtitle("Imputed only - logreg")
marrangeGrob(g,nrow=1,ncol=4,top="Covariance")
#marrangeGrob(g,nrow=2,ncol=2,top="Covariance")

vars = c(fpVar,sprintf("%s_prev",fpVar))
g = list()
g[[1]] = TilePlot(cov(long[,vars],use='pairwise.complete'),range=xlim)+ggtitle("Available-case")
g[[2]] = TilePlot(Cv[[1]][vars,vars],dm=Cv[[2]][vars,vars],range=xlim)+ggtitle("Imputed - logreg")
g[[3]] = TilePlot(Cvi[[1]][vars,vars],dm=Cvi[[2]][vars,vars],range=xlim)+ggtitle("Imputed only - logreg")
marrangeGrob(g,nrow=1,ncol=3,top="Covariance")
#marrangeGrob(g,nrow=2,ncol=2,top="Covariance")
```

```{r}
library(mgcv)
gm =  list()
testage = seq(min(long[,"age"],na.rm=T)-1,max(long[,"age"],na.rm=T)+1,lenght=101)
for (j in 1:length(agg_fpvar))
{
  fitdata = subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0 & x > 0 & x_se > 0)
  #mexp = lm(log(x)~age_cut,fitdata,weights=1/fitdata[,"x_se"]^2)
  #g = gam(x~s(age_cut,3),data=fitdata,weights=1/fitdata[,"x_se"]^2)
  
  Nimp = sum(ml_fp$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gm[[j]] = gm[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth(data=subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0),mapping=aes(x=age_cut,y=x),inherit.aes=F,se=F,lty=2,method="gam",formula=y~s(x,4))+
  #geom_line(data=data.frame(age=testage,y=exp(predict(mexp,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  #geom_line(data=data.frame(age=testage,y=exp(predict(g,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1,fullrange=T)+
  labs(x="Age",title=sprintf("%s (males)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}

gm[[length(gm)+1]] = ggplot(data=subset(agg,sex==0 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gm,nrow=2,ncol=3,top="males")

gf =  list()
for (j in 1:length(agg_fpvar))
{
    fitdata = subset(agg_fpvar[[j]], N > 1 & sex==1 & imp == 0 & x > 0 & x_se > 0)
  Nimp = sum(ml_fp$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gf[[j]] = gf[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth()+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1)+
  labs(x="Age",title=sprintf("%s (females)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}


#females
gf[[length(gf)+1]] = ggplot(data=subset(agg,sex==1 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  #scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gf,nrow=2,ncol=3,top="females")


save=T
if(save)
{   
    ggsave(sprintf("%s/results/hrs_mice_lm_males_fp.pdf",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/hrs_mice_lm_males_fp.png",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    
      ggsave(sprintf("%s/results/hrs_mice_lm_females_fp.pdf",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/hrs_mice_lm_females_fp.png",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
}
```


# to do: check FI


these ones are better
```{r}
library(colorspace)

#pointless, so few options for values
#ggplot(data=subset(fp,sex=="male"),aes(x=Nfp_noimp,y=Nfp,color=ordered(Nfp_imputed)))+
#  geom_point()+
  #geom_smooth(se=F)+
#  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
#  theme_minimal()





#combined sexes #doesn't include imputatoin error
#ggplot(data=fp,aes(x=round(age,0),y=Nfp,color=ordered(N_imputed)))+
#  stat_summary()+
#  geom_smooth(se=F)+
#  scale_y_log10()+
#  annotation_logticks(sides="l")+
#  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
#  theme_minimal()

#males
#note: errorbars aren't right for N=1 since sd -> NA gets ignored
ggplot(data=subset(agg,sex==0 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

#females
ggplot(data=subset(agg,sex==1 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

ggplot(data=subset(agg, N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp),shape=ordered(sex)),linetype=ordered(sex))+
  geom_pointrange()+
  #geom_smooth(se=F)+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

ggplot(data=subset(agg, N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp),shape=ordered(sex)),linetype=ordered(sex))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=Nfp,weight=1/Nfp_se^2),lty=2,inherit.aes=F,fullrange =T)+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()



```

```{r}
library(colorspace)
fi = list()
for (i in 1:ml_fi$m)
{
  fi[[i]] = mice::complete(ml_fi,i)[,c("id","age","sex","wave",fiVar)]
  fi[[i]][,"imp"] = i
  fi[[i]][,"fi"] = apply(fi[[i]][,fiVar],1,mean,na.rm=T)
  fi[[i]][,"fi_noimp"] = apply(long[,fiVar],1,mean,na.rm=T)
  fi[[i]][,sprintf("%s_imputed",fiVar)] = ml_fp$where[,fiVar]
  fi[[i]][,"N_imputed"] = apply(ml_fp$where[,fiVar],1,sum)
}
fi = do.call(rbind,fi)
age_cuts = seq(min(fi[,"age"])-1,max(fi[,"age"])+1,length=8)
fi[,"age_cut"] = cut(fi[,"age"],age_cuts,include.lowest=T)

ggplot(data=subset(fi,sex=="male"),aes(x=round(age,0),y=fi,color=ordered(N_imputed)))+
  stat_summary()+
  geom_smooth(se=F)+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

ggplot(data=subset(fi,sex=="female"),aes(x=round(age,0),y=fi,color=ordered(N_imputed)))+
  stat_summary()+
  geom_smooth(se=F)+
  scale_y_log10()+
  annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
```

split cart
```{r}
save = T
load = T
file = sprintf("%s/data/hrs_rand_fi_vs_fp_mice_cart_fp.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  mcart_fp = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),"age","sex","wave","id")
  mcart_fp = mice(long[,vars], meth = methCart[vars],pred=predCart[vars,vars], m = 15, print = T)
  if(save) saveRDS(mcart_fp,file)
}


file = sprintf("%s/data/hrs_rand_fi_vs_fp_mice_cart_fi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  mcart_fi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id")
  mcart_fi = mice(long[,vars], meth = methCart[vars],pred=predCart[vars,vars], m = 15, print = T)
  if(save) saveRDS(mcart_fi,file)
}


file = sprintf("%s/data/hrs_rand_fi_vs_fp_mice_cart_fpfi_markov.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  mcart_fpfi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id") #crashing
  #vars = c(fpVar,sprintf("%s_prev",fpVar),fiVar,sprintf("%s_prev",fiVar),"age","sex","id") #also crashing
  me = methCart[vars]
  pr = predCart[vars,vars]
  #pr[grep("prev",names(me)),] = 0 #don't impute back in time
  #pr[grep("age",names(me)),] = 1
  #pr[grep("sex",names(me)),] = 1
  #pr[grep("wave",names(me)),] = 1
  #me[grep("prev",names(me))] = "" #don't impute prev #pretty sure this won't work
  mcart_fpfi = mice(long[,vars], meth = me,pred=pr, m = 15, print = T)
  if(save) saveRDS(mcart_fpfi,file)
}

file = sprintf("%s/data/hrs_rand_fi_vs_fp_mice_cart_fpfi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  mcart_fpfi2 = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,fiVar,"age","sex","wave","id")
  me = methCart[vars]
  pr = predCart[vars,vars]
  mcart_fpfi2 = mice(long[,vars], meth = me,pred=pr, m = 15, print = T)
  if(save) saveRDS(mcart_fpfi2,file)
}
```

```{r}
mmice = mcart_fpfi2
library(colorspace)
fp = list()
fi = list()
C = list() #correlation
Cv = list() #covariance
Cvi = list() #covariance - imputed values only
mu = list()
age_cuts = seq(min(long[,"age"])-1,max(long[,"age"])+1,length=10)
imp_cuts = seq(0,100,length=5) #for FI
agg = list()
aggse = list()
aggfi = list()
aggfise = list()
agg_fpvar = list() #one per fpvar
agg_fpvar_se = list()
for (j in 1:length(fpVar)) 
{
  agg_fpvar[[j]] = list()
  agg_fpvar_se[[j]]  = list()
}
for (i in 1:mmice$m)
{
  fp[[i]] = mice::complete(mmice,i)[,c("id","age","sex","wave",fpVar)]
  fp[[i]][,"imp"] = i
  fp[[i]][,"Nfp"] = apply(fp[[i]][,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,"Nfp_noimp"] = apply(long[,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,sprintf("%s_imputed",fpVar)] = mmice$where[,fpVar]
  fp[[i]][,"N_imputed"] = apply(mmice$where[,fpVar],1,sum)
  C[[i]] = cor(mice::complete(mmice,i)[,c("age",fpVar)])
  Cv[[i]] = cov(mice::complete(mmice,i)[,c("age",fpVar)])
  temp = mice::complete(mmice,i)[,c(fpVar)]
  temp[!mmice$where[,c(fpVar)]] = NA
  Cvi[[i]] = cov(temp,use='pairwise.complete')
  mu[[i]] = apply(mice::complete(mmice,i)[,c("age",fpVar)],2,mean,na.rm=T)
  m = matrix(NA,nrow=length(mu[[i]]),ncol=2)
  m[,1] = mu[[i]]
  m[,2] = mu[[i]]
  mu[[i]] = m
  
  
  fp[[i]][,"age_cut"] = as.numeric(as.character(cut(fp[[i]][,"age"],age_cuts,labels=age_cuts[-1]/2+age_cuts[-length(age_cuts)]/2,include.lowest=T)))
  
  vars = c("age",fpVar,"Nfp","Nfp_noimp")
  agg[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
  aggse[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
  agg[[i]][,"N"] = aggregate(fp[[i]][,"age"],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
  agg[[i]][,"sex"] = as.numeric(factor(agg[[i]][,"sex"]),c("male","female"))-1
  aggse[[i]][,"N"] = 0
  aggse[[i]][,"sex"] = 0

  for(j in 1:length(fpVar))
  {
    agg_fpvar[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=mmice$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
    agg_fpvar_se[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=mmice$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
    agg_fpvar[[j]][[i]][,"N"] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=mmice$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
    agg_fpvar[[j]][[i]][,"sex"] = as.numeric(factor(agg_fpvar[[j]][[i]][,"sex"]),c("male","female"))-1
    agg_fpvar_se[[j]][[i]][,"N"] = 0
    agg_fpvar_se[[j]][[i]][,"sex"] = 0
  }
  
  
  fi[[i]] = mice::complete(mmice,i)[,c("id","age","sex","wave",fiVar)]
  fi[[i]][,"imp"] = i
  fi[[i]][,"fi"] = apply(fi[[i]][,fiVar],1,mean,na.rm=T)
  fi[[i]][,"fi_noimp"] = apply(long[,fiVar],1,mean,na.rm=T)
  fi[[i]][,sprintf("%s_imputed",fiVar)] = mmice$where[,fiVar]
  fi[[i]][,"N_imputed"] = apply(mmice$where[,fiVar],1,sum)
  
  fi[[i]][,"age_cut"] = as.numeric(as.character(cut(fi[[i]][,"age"],age_cuts,labels=age_cuts[-1]/2+age_cuts[-length(age_cuts)]/2,include.lowest=T)))
  fi[[i]][,"percent_imp_cut"] = as.numeric(as.character(cut(fi[[i]][,"N_imputed"]/length(fiVar)*100,imp_cuts,labels=round(imp_cuts[-1]/2+imp_cuts[-length(imp_cuts)]/2),include.lowest=T)))
  aggfi[[i]] = aggregate(fi[[i]][,c("fi","fi_noimp","N_imputed")],by=list(sex=fi[[i]][,"sex"],percent_imp_cut=fi[[i]][,"percent_imp_cut"],age_cut=fi[[i]][,"age_cut"]),mean,na.rm=T)
  aggfise[[i]] = aggregate(fi[[i]][,c("fi","fi_noimp","N_imputed")],by=list(sex=fi[[i]][,"sex"],percent_imp_cut=fi[[i]][,"percent_imp_cut"],age_cut=fi[[i]][,"age_cut"]),SEM,na.rm=T)
  aggfi[[i]][,"N"] = aggregate(fi[[i]][,"age"],by=list(sex=fi[[i]][,"sex"],percent_imp_cut=fi[[i]][,"percent_imp_cut"],age_cut=fi[[i]][,"age_cut"]),length)[,"x"]
  aggfi[[i]][,"sex"] = as.numeric(factor(aggfi[[i]][,"sex"]),c("male","female"))-1
  aggfise[[i]][,"N"] = 0
  aggfise[[i]][,"sex"] = 0

}
fp = do.call(rbind,fp)
fi = do.call(rbind,fi)
C = RubinMat(C)
Cv = RubinMat(Cv)
Cvi = RubinMat(Cvi)
mu = RubinMat(mu)

agg = RubinMat(agg,lse=aggse)
vars = c(fpVar,"Nfp","Nfp_noimp")
agg[[1]] = as.data.frame(agg[[1]])
agg[[1]][,sprintf("%s_se",vars)] = agg[[2]][,vars]
agg = agg[[1]]

aggfi = RubinMat(aggfi,lse=aggfise)
vars = c("fi","fi_noimp")
aggfi[[1]] = as.data.frame(aggfi[[1]])
aggfi[[1]][,sprintf("%s_se",vars)] = aggfi[[2]][,vars]
aggfi = aggfi[[1]]

for (j in 1:length(fpVar))
{
  agg_fpvar[[j]] = RubinMat(agg_fpvar[[j]],agg_fpvar_se[[j]])
  agg_fpvar[[j]][[1]] = as.data.frame(agg_fpvar[[j]][[1]])
  agg_fpvar[[j]][[1]][,sprintf("%s_se","x")] = agg_fpvar[[j]][[2]][,"x"]
  agg_fpvar[[j]] = agg_fpvar[[j]][[1]]
}



```
```{r}
#gm = ggplot(data=subset(aggfi,sex==0 & N > 1),aes(x=age_cut,y=fi,ymin=fi-fi_se,ymax=fi+fi_se,color=ordered(Nimp)))+
gm = ggplot(data=subset(aggfi,sex==0 & N > 1),aes(x=age_cut,y=fi,ymin=fi-fi_se,ymax=fi+fi_se,color=ordered(percent_imp_cut)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  ggtitle("Males")+
  theme_minimal()
gm

plotdata = list()
plotdata[[1]] = subset(aggfi,sex==0 & N > 1)
plotdata[[1]][,"group"] = "mice"
plotdata[[2]] = subset(aggfi,sex==0 & N > 1)
plotdata[[2]][,"fi"] = plotdata[[2]][,"fi_noimp"]
plotdata[[2]][,"fi_se"] = plotdata[[2]][,"fi_noimp_se"]
plotdata[[2]][,"group"] = "ignore"
plotdata = do.call(rbind,plotdata)

gm2 = ggplot(data=plotdata,aes(x=age_cut,y=fi,ymin=fi-fi_se,ymax=fi+fi_se,color=ordered(group)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  ggtitle("Males")+
  theme_minimal()

#gm2


#gf = ggplot(data=subset(aggfi,sex==1 & N > 1),aes(x=age_cut,y=fi,ymin=fi-fi_se,ymax=fi+fi_se,color=ordered(Nimp)))+
gf = ggplot(data=subset(aggfi,sex==1 & N > 1),aes(x=age_cut,y=fi,ymin=fi-fi_se,ymax=fi+fi_se,color=ordered(percent_imp_cut)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  ggtitle("Females")+
  theme_minimal()
gf

plotdata = list()
plotdata[[1]] = subset(aggfi,sex==1 & N > 1)
plotdata[[1]][,"group"] = "mice"
plotdata[[2]] = subset(aggfi,sex==1 & N > 1)
plotdata[[2]][,"fi"] = plotdata[[2]][,"fi_noimp"]
plotdata[[2]][,"fi_se"] = plotdata[[2]][,"fi_noimp_se"]
plotdata[[2]][,"group"] = "ignore"
plotdata = do.call(rbind,plotdata)

gf2 = ggplot(data=plotdata,aes(x=age_cut,y=fi,ymin=fi-fi_se,ymax=fi+fi_se,color=ordered(group)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

#gf2

save=T
if(save)
{   
    ggsave(sprintf("%s/results/hrs_mice_cart_fi.pdf",outputDir),
           marrangeGrob(list(gm,gf),nrow=1,ncol=2,top=NULL),width=16,height=7,dpi=300)
    ggsave(sprintf("%s/results/hrs_mice_cart_fi.png",outputDir),
           marrangeGrob(list(gm,gf),nrow=2,ncol=2,top=NULL),width=16,height=7,dpi=300)
}
```

```{r}
library(mgcv)
gm =  list()
testage = seq(min(long[,"age"],na.rm=T)-1,max(long[,"age"],na.rm=T)+1,lenght=101)
for (j in 1:length(agg_fpvar))
{
  fitdata = subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0 & x > 0 & x_se > 0)
  #mexp = lm(log(x)~age_cut,fitdata,weights=1/fitdata[,"x_se"]^2)
  #g = gam(x~s(age_cut,3),data=fitdata,weights=1/fitdata[,"x_se"]^2)
  
  Nimp = sum(mmice$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gm[[j]] = gm[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth(data=subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0),mapping=aes(x=age_cut,y=x),inherit.aes=F,se=F,lty=2,method="gam",formula=y~s(x,4))+
  #geom_line(data=data.frame(age=testage,y=exp(predict(mexp,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  #geom_line(data=data.frame(age=testage,y=exp(predict(g,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1,fullrange=T)+
  labs(x="Age",title=sprintf("%s (males)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}

gm[[length(gm)+1]] = ggplot(data=subset(agg,sex==0 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gm,nrow=2,ncol=3,top="males")

gf =  list()
for (j in 1:length(agg_fpvar))
{
    fitdata = subset(agg_fpvar[[j]], N > 1 & sex==1 & imp == 0 & x > 0 & x_se > 0)
  Nimp = sum(mmice$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gf[[j]] = gf[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth()+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1)+
  labs(x="Age",title=sprintf("%s (females)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}


#females
gf[[length(gf)+1]] = ggplot(data=subset(agg,sex==1 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  #scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gf,nrow=2,ncol=3,top="females")


save=T
if(save)
{   
    ggsave(sprintf("%s/results/hrs_mice_cart_males_fp.pdf",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/hrs_mice_cart_males_fp.png",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    
      ggsave(sprintf("%s/results/hrs_mice_cart_females_fp.pdf",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/hrs_mice_cart_females_fp.png",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
}
```

pooled cart
```{r}
save = T
load = T
file = sprintf("%s/data/hrs_fi_vs_fp_mice_cart.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  mcart = readRDS(file)
} else
{
  print("file not found, computing...")
  #mcart = mice(long[,setdiff(colnames(long),"wave")], meth = "cart", m = 1, maxit= 1, print = T)
  mcart = mice(long, meth = methCart,pred=predCart, m = 15, print = T)
  if(save) saveRDS(mcart,file)
}

```



#
# Missingness statistics
#
```{r}
plotdata = data.frame(missingness = apply(is.na(long),2,mean),names=colnames(long),
                      N = nrow(long)
                      )
#not sure what error is...
#https://www.researchgate.net/post/Can-standard-deviation-and-standard-error-be-calculated-for-a-binary-variable
#sqrt(p.est*(1-p.est)/n)
plotdata[,"se"] = sqrt(plotdata[,"missingness"]*(1-plotdata[,"missingness"])/plotdata[,"N"])
plotdata[,"names"] = factor(plotdata[,"names"],plotdata[sort.list(plotdata[,"missingness"],decreasing=T),"names"])
ggplot(plotdata,aes(x=names,y=missingness,ymin=missingness-se,ymax=missingness+se))+
  geom_pointrange()+
  labs(x="")+
  theme_minimal()+
  theme(axis.text.x=element_text(angle=90,vjust=.5))
```

Just a few variables dominate the missingness, namely weight and gait speed.

```{r}
#relation between missingness and health
plotdata = data.frame(missing = apply(is.na(long),1,sum))

plotdata[,"fi"] = apply(long[,fiVar],1,mean,na.rm=T)
plotdata[,"age"] = long[,"age"]
plotdata[,"fic"] = cut(plotdata[,"fi"],c(-Inf,.1,.21,.45,Inf)) #blodgett cuts
plotdata[,"agec"] = cut(plotdata[,"age"],c(0,60,70,80,Inf))
ggplot(plotdata,aes(x=fi,y=missing,colour=agec,shape=agec))+geom_point()+geom_smooth()
```

No obvious connection between FI, age and missingness.

```{r}
#relation between missingness and health
plotdata = data.frame(missingNFP = apply(is.na(long[,fpVar]),1,sum))

plotdata[,"fi"] = apply(long[,fiVar],1,mean,na.rm=T)
plotdata[,"age"] = long[,"age"]
plotdata[,"fic"] = cut(plotdata[,"fi"],c(-Inf,.1,.21,.45,Inf)) #blodgett cuts
plotdata[,"agec"] = cut(plotdata[,"age"],c(0,60,70,80,Inf))
ggplot(plotdata,aes(x=age,y=missingNFP,colour=fic,shape=fic))+stat_summary()+geom_smooth()
```

```{r}
stop("depricated code")
```

```{r}
gm =  list()
gm[[1]] = ggplot(data=subset(agg, N > 1 & sex==0),aes(x=age_cut,y=fp1_weight,ymin=fp1_weight-fp1_weight_se,ymax=fp1_weight+fp1_weight_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp1_weight,weight=1/fp1_weight_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp1 weight (males)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gm[[2]] = ggplot(data=subset(agg, N > 1 & sex==0),aes(x=age_cut,y=fp2_grip,ymin=fp2_grip-fp2_grip_se,ymax=fp2_grip+fp2_grip_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp2_grip,weight=1/fp2_grip_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp2 grip (males)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gm[[3]] = ggplot(data=subset(agg, N > 1 & sex==0),aes(x=age_cut,y=fp3_gait,ymin=fp3_gait-fp3_gait_se,ymax=fp3_gait+fp3_gait_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp3_gait,weight=1/fp3_gait_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp3 gait (males)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gm[[4]] = ggplot(data=subset(agg, N > 1 & sex==0),aes(x=age_cut,y=fp4_exhaustion,ymin=fp4_exhaustion-fp4_exhaustion_se,ymax=fp4_exhaustion+fp4_exhaustion_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp4_exhaustion,weight=1/fp4_exhaustion_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp4 exhaustion (males)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gm[[5]] = ggplot(data=subset(agg, N > 1 & sex==0),aes(x=age_cut,y=fp5_low_activity,ymin=fp5_low_activity-fp5_low_activity_se,ymax=fp5_low_activity+fp5_low_activity_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp5_low_activity,weight=1/fp5_low_activity_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp5 low activity (males)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

marrangeGrob(gm,nrow=2,ncol=3,top="males")

gf =  list()
gf[[1]] = ggplot(data=subset(agg, N > 1 & sex==1),aes(x=age_cut,y=fp1_weight,ymin=fp1_weight-fp1_weight_se,ymax=fp1_weight+fp1_weight_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp1_weight,weight=1/fp1_weight_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp1 weight (females)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gf[[2]] = ggplot(data=subset(agg, N > 1 & sex==1),aes(x=age_cut,y=fp2_grip,ymin=fp2_grip-fp2_grip_se,ymax=fp2_grip+fp2_grip_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp2_grip,weight=1/fp2_grip_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp2 grip (females)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gf[[3]] = ggplot(data=subset(agg, N > 1 & sex==1),aes(x=age_cut,y=fp3_gait,ymin=fp3_gait-fp3_gait_se,ymax=fp3_gait+fp3_gait_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp3_gait,weight=1/fp3_gait_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp3 gait (females)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gf[[4]] = ggplot(data=subset(agg, N > 1 & sex==1),aes(x=age_cut,y=fp4_exhaustion,ymin=fp4_exhaustion-fp4_exhaustion_se,ymax=fp4_exhaustion+fp4_exhaustion_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp4_exhaustion,weight=1/fp4_exhaustion_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp4 exhaustion (females)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

gf[[5]] = ggplot(data=subset(agg, N > 1 & sex==1),aes(x=age_cut,y=fp5_low_activity,ymin=fp5_low_activity-fp5_low_activity_se,ymax=fp5_low_activity+fp5_low_activity_se,color=ordered(Nimp)))+
  geom_pointrange(position=position_dodge(1))+
  geom_smooth(method="lm",data=subset(agg,N>1 & Nimp==0),aes(x=age_cut,y=fp5_low_activity,weight=1/fp5_low_activity_se^2),lty=2,inherit.aes=F,fullrange =T)+
  labs(x="Age",title="fp5 low activity (females)")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()

marrangeGrob(gf,nrow=2,ncol=3,top="females")
```

