---
title: "Imputation - ELSA"
output: html_notebook
---

Apr 2024
Updated imputation method. Not sure if better than old method, will check (old method is in fi_vs_fp_imputation)

MICE. Algorithm(s): 
-causal CART
-logistic/normal regression

```{r}
outputDir = "/home/glen/analysis/fi_vs_fp"
gRootDir = "/home/glen/Documents/r" #where scripts are

setwd(outputDir)
print(getwd())

source(sprintf("%s/nhanes.R",gRootDir),verbose=0)
source(sprintf("%s/goodness_of_fit.R",gRootDir),verbose=0)
source(sprintf("%s/pca.R",gRootDir),verbose=0)
source(sprintf("%s/pca_fi.R",gRootDir),verbose=0)
library("mice")
```
#(almost) everybody has at least one: grip or gait
```{r}
elsa = readRDS(sprintf("%s/data/elsa_fi_vs_fp.rds",outputDir))
```

```{r}
# narrow down variables
adlVar = c("mob_walk100yards",
"mob_sit2hours",
"mob_situp",
"mob_climbflights",
"mob_climb1flight",
"mob_stoop",
"mob_armsup",
"mob_pushpull",
"mob_liftcarry",
"pick_up_coin",
"dress",
"adl_walkroom",
"adl_bathe",
"adl_eat",
"adl_bed",
"adl_toilet",
"iadl_map",
"iadl_makemeal",
"iadl_groceries",
"iadl_phone",
"iadl_meds",
"iadl_housework",
"iadl_money"
)

genVar = c('general','eye','hear')

lungVar = c('lung_disease',
           'asthma',
           'arthritis',
           'osteoporosis',
           'cancer',
           'parkinsons',
           'psychiatric_problem',
           'alzheimers',
           'dementia'
            )

heartVar = c('hyptertension',
           'angina',
           'heart_attack',
           'heart_failure',
           'heart_murmur',
           'abnormal_heart',
           'diabetes',
           'stroke'
            )

disVar = c(lungVar,heartVar)

fiVar = c(adlVar,genVar,disVar)

fpVar = c('fp1_weight','fp2_grip','fp3_gait','fp4_exhaustion','fp5_low_activity')

auxVar = c("age","sex") #"walk_time_ave","height","weight" # I'm worried about overfitting #"fi_prev" probably good but I'm concerned I'll get pushback
```

#
# long format imputation  - without instantiating new timepoints
#

# convert to long format
```{r}
long = list()
for (i in 1:length(elsa))
{
  long[[i]] = elsa[[i]][,c("id",fiVar,fpVar,auxVar)]
  long[[i]][,"wave"] = names(elsa)[i] #easier to decode afterwards
  
  long[[i]][,sprintf("%s_prev",fiVar)] = NA
  long[[i]][,sprintf("%s_prev",fpVar)] = NA
  long[[i]][,"fi_prev"] = NA
  long[[i]][,"Nfp_prev"] = NA
  if(i> 1)
  {
    ids = intersect(rownames(long[[i-1]]),rownames(long[[i]]))

    long[[i]][ids,sprintf("%s_prev",fiVar)] = long[[i-1]][ids,fiVar]
    long[[i]][ids,"fi_prev"] =          apply(long[[i-1]][ids,fiVar],1,mean,na.rm=T)
    availfpVar = intersect(colnames(long[[i-1]]),fpVar)
    print(availfpVar)
    long[[i]][ids,sprintf("%s_prev",availfpVar)] = long[[i-1]][ids,availfpVar]
    long[[i]][ids,"Nfp_prev"] =          apply(long[[i-1]][ids,availfpVar],1,mean,na.rm=T)*length(availfpVar)
  }
}
long = do.call(rbind,long)
```

```{r}
long[,"id"] = as.integer(long[,"id"]) #mice wants integer here or it gets mad

long = long[,c("id","wave","age","sex",fiVar,fpVar,sprintf("%s_prev",fiVar),sprintf("%s_prev",fpVar))]

#'fix' types
for (j in 1:length(fiVar))
{
  long[,fiVar[j]] = as.integer(long[,fiVar[j]])
  long[,sprintf("%s_prev",fiVar[j])] = as.integer(long[,sprintf("%s_prev",fiVar[j])])
  #long[,fiVar[j]] = ordered(long[,fiVar[j]])
  #long[,sprintf("%s_prev",fiVar[j])] = ordered(long[,sprintf("%s_prev",fiVar[j])])
}
for (j in 1:length(fpVar))
{
  long[,fpVar[j]] = as.integer(long[,fpVar[j]])
  long[,sprintf("%s_prev",fpVar[j])] = as.integer(long[,sprintf("%s_prev",fpVar[j])])
  #long[,fpVar[j]] = ordered(long[,fpVar[j]])
  #long[,sprintf("%s_prev",fpVar[j])] = ordered(long[,sprintf("%s_prev",fpVar[j])])
}
```

```{r}
ordImp = c(genVar,sprintf("%s_prev",genVar))
binImp = setdiff(c(fiVar,fpVar,sprintf("%s_prev",c(fiVar,fpVar))),ordImp)

pred = make.predictorMatrix(long)
pred[,"wave"] = 0 #don't predict wave


#basic linear models
meth = make.method(long)
meth[1:length(meth)] = ""
meth[binImp] = "logreg"
meth[ordImp] = "polr" #"2l.norm" #"2l.norm" #2l.pan doesn't allow missing values in predictors
print(meth)
```

```{r}
predCart = make.predictorMatrix(long)

methCart = make.method(long,defaultMethod="cart")
methCart["wave"] = ""
methCart[c(binImp,ordImp)] = "cart"

print(methCart) #some may have no NAs in which case they default to ""
```

split linear
```{r}
save = T
load = T
file = sprintf("%s/data/elsa_fi_vs_fp_mice_lin_fp.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_fp = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),"age","sex","wave","id")
  ml_fp = mice(long[,vars], meth = meth[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_fp,file)
}


file = sprintf("%s/data/elsa_fi_vs_fp_mice_lin_fi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_fi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id")
  ml_fi = mice(long[,vars], meth = meth[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_fi,file)
}


file = sprintf("%s/data/elsa_fi_vs_fp_mice_lin_fpfi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_fpfi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id")
  ml_fpfi = mice(long[,vars], meth = meth[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_fpfi,file)
}
```

fit summary - fp variables
```{r}
library(colorspace)
fp = list()
C = list() #correlation
Cv = list() #covariance
Cvi = list() #covariance - imputed values only
mu = list()
age_cuts = seq(min(long[,"age"])-1,max(long[,"age"])+1,length=10)
agg = list()
aggse = list()
agg_fpvar = list() #one per fpvar
agg_fpvar_se = list()
for (j in 1:length(fpVar)) 
{
  agg_fpvar[[j]] = list()
  agg_fpvar_se[[j]]  = list()
}
for (i in 1:ml_fp$m)
{
  fp[[i]] = mice::complete(ml_fp,i)[,c("id","age","sex","wave",fpVar)]
  fp[[i]][,"imp"] = i
  fp[[i]][,"Nfp"] = apply(fp[[i]][,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,"Nfp_noimp"] = apply(long[,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,sprintf("%s_imputed",fpVar)] = ml_fp$where[,fpVar]
  fp[[i]][,"N_imputed"] = apply(ml_fp$where[,fpVar],1,sum)
  C[[i]] = cor(mice::complete(ml_fp,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))])
  Cv[[i]] = cov(mice::complete(ml_fp,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))])
  temp = mice::complete(ml_fp,i)[,c(fpVar,sprintf("%s_prev",fpVar))]
  temp[!ml_fp$where[,c(fpVar,sprintf("%s_prev",fpVar))]] = NA
  Cvi[[i]] = cov(temp,use='pairwise.complete')
  mu[[i]] = apply(mice::complete(ml_fp,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))],2,mean,na.rm=T)
  m = matrix(NA,nrow=length(mu[[i]]),ncol=2)
  m[,1] = mu[[i]]
  m[,2] = mu[[i]]
  mu[[i]] = m
  
  
  fp[[i]][,"age_cut"] = as.numeric(as.character(cut(fp[[i]][,"age"],age_cuts,labels=age_cuts[-1]/2+age_cuts[-length(age_cuts)]/2,include.lowest=T)))
  
  vars = c("age",fpVar,"Nfp","Nfp_noimp")
  agg[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
  aggse[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
  agg[[i]][,"N"] = aggregate(fp[[i]][,"age"],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
  agg[[i]][,"sex"] = as.numeric(factor(agg[[i]][,"sex"]),c("male","female"))-1
  aggse[[i]][,"N"] = 0
  aggse[[i]][,"sex"] = 0

  for(j in 1:length(fpVar))
  {
    agg_fpvar[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml_fp$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
    agg_fpvar_se[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml_fp$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
    agg_fpvar[[j]][[i]][,"N"] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml_fp$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
    agg_fpvar[[j]][[i]][,"sex"] = as.numeric(factor(agg_fpvar[[j]][[i]][,"sex"]),c("male","female"))-1
    agg_fpvar_se[[j]][[i]][,"N"] = 0
    agg_fpvar_se[[j]][[i]][,"sex"] = 0
  }

}
fp = do.call(rbind,fp)
C = RubinMat(C)
Cv = RubinMat(Cv)
Cvi = RubinMat(Cvi)
mu = RubinMat(mu)

agg = RubinMat(agg,lse=aggse)
vars = c(fpVar,"Nfp","Nfp_noimp")
agg[[1]] = as.data.frame(agg[[1]])
agg[[1]][,sprintf("%s_se",vars)] = agg[[2]][,vars]
agg = agg[[1]]

for (j in 1:length(fpVar))
{
  agg_fpvar[[j]] = RubinMat(agg_fpvar[[j]],agg_fpvar_se[[j]])
  agg_fpvar[[j]][[1]] = as.data.frame(agg_fpvar[[j]][[1]])
  agg_fpvar[[j]][[1]][,sprintf("%s_se","x")] = agg_fpvar[[j]][[2]][,"x"]
  agg_fpvar[[j]] = agg_fpvar[[j]][[1]]
}



```

fit summary - fp variables
```{r}
library(mgcv)
gm =  list()
testage = seq(min(long[,"age"],na.rm=T)-1,max(long[,"age"],na.rm=T)+1,lenght=101)
for (j in 1:length(agg_fpvar))
{
  fitdata = subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0 & x > 0 & x_se > 0)
  #mexp = lm(log(x)~age_cut,fitdata,weights=1/fitdata[,"x_se"]^2)
  #g = gam(x~s(age_cut,3),data=fitdata,weights=1/fitdata[,"x_se"]^2)
  
  Nimp = sum(ml_fp$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gm[[j]] = gm[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth(data=subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0),mapping=aes(x=age_cut,y=x),inherit.aes=F,se=F,lty=2,method="gam",formula=y~s(x,4))+
  #geom_line(data=data.frame(age=testage,y=exp(predict(mexp,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  #geom_line(data=data.frame(age=testage,y=exp(predict(g,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1,fullrange=T)+
  labs(x="Age",title=sprintf("%s (males)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}

gm[[length(gm)+1]] = ggplot(data=subset(agg,sex==0 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gm,nrow=2,ncol=3,top="males")

gf =  list()
for (j in 1:length(agg_fpvar))
{
    fitdata = subset(agg_fpvar[[j]], N > 1 & sex==1 & imp == 0 & x > 0 & x_se > 0)
  Nimp = sum(ml_fp$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gf[[j]] = gf[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth()+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1)+
  labs(x="Age",title=sprintf("%s (females)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}


#females
gf[[length(gf)+1]] = ggplot(data=subset(agg,sex==1 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  #scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gf,nrow=2,ncol=3,top="females")


save=T
if(save)
{   
    ggsave(sprintf("%s/results/elsa_lm_males_fp.pdf",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/elsa_lm_males_fp.png",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    
      ggsave(sprintf("%s/results/elsa_lm_females_fp.pdf",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/elsa_lm_females_fp.png",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
}
```

split cart
```{r}
save = T
load = T
file = sprintf("%s/data/elsa_fi_vs_fp_mice_cart_fp.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_cart_fp = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),"age","sex","wave","id")
  ml_cart_fp = mice(long[,vars], meth = methCart[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_cart_fp,file)
}


file = sprintf("%s/data/elsa_fi_vs_fp_mice_cart_fi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_cart_fi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id")
  ml_cart_fi = mice(long[,vars], meth = methCart[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_cart_fi,file)
}

file = sprintf("%s/data/elsa_fi_vs_fp_mice_cart_fpfi.rds",outputDir)
if(file.exists(file) & load)
{
  print("file found, loading...")
  ml_cart_fpfi = readRDS(file)
} else
{
  print("file not found, computing...")
  vars = c(fpVar,sprintf("%s_prev",fpVar),fiVar,sprintf("%s_prev",fiVar),"age","sex","wave","id")
  ml_cart_fpfi = mice(long[,vars], meth = methCart[vars],pred=pred[vars,vars], m = 15, print = T)
  if(save) saveRDS(ml_cart_fpfi,file)
}
```

check fit

fit summary - fp variables
```{r}
library(colorspace)
ml = ml_cart_fpfi
fp = list()
C = list() #correlation
Cv = list() #covariance
Cvi = list() #covariance - imputed values only
mu = list()
age_cuts = seq(min(long[,"age"])-1,max(long[,"age"])+1,length=10)
agg = list()
aggse = list()
agg_fpvar = list() #one per fpvar
agg_fpvar_se = list()
for (j in 1:length(fpVar)) 
{
  agg_fpvar[[j]] = list()
  agg_fpvar_se[[j]]  = list()
}
for (i in 1:ml$m)
{
  fp[[i]] = mice::complete(ml,i)[,c("id","age","sex","wave",fpVar)]
  fp[[i]][,"imp"] = i
  fp[[i]][,"Nfp"] = apply(fp[[i]][,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,"Nfp_noimp"] = apply(long[,fpVar],1,mean,na.rm=T)*length(fpVar)
  fp[[i]][,sprintf("%s_imputed",fpVar)] = ml$where[,fpVar]
  fp[[i]][,"N_imputed"] = apply(ml$where[,fpVar],1,sum)
  C[[i]] = cor(mice::complete(ml,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))])
  Cv[[i]] = cov(mice::complete(ml,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))])
  temp = mice::complete(ml,i)[,c(fpVar,sprintf("%s_prev",fpVar))]
  temp[!ml$where[,c(fpVar,sprintf("%s_prev",fpVar))]] = NA
  Cvi[[i]] = cov(temp,use='pairwise.complete')
  mu[[i]] = apply(mice::complete(ml,i)[,c("age",fpVar,sprintf("%s_prev",fpVar))],2,mean,na.rm=T)
  m = matrix(NA,nrow=length(mu[[i]]),ncol=2)
  m[,1] = mu[[i]]
  m[,2] = mu[[i]]
  mu[[i]] = m
  
  
  fp[[i]][,"age_cut"] = as.numeric(as.character(cut(fp[[i]][,"age"],age_cuts,labels=age_cuts[-1]/2+age_cuts[-length(age_cuts)]/2,include.lowest=T)))
  
  vars = c("age",fpVar,"Nfp","Nfp_noimp")
  agg[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
  aggse[[i]] = aggregate(fp[[i]][,vars],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
  agg[[i]][,"N"] = aggregate(fp[[i]][,"age"],by=list(sex=fp[[i]][,"sex"],Nimp=fp[[i]][,"N_imputed"],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
  agg[[i]][,"sex"] = as.numeric(factor(agg[[i]][,"sex"]),c("male","female"))-1
  aggse[[i]][,"N"] = 0
  aggse[[i]][,"sex"] = 0

  for(j in 1:length(fpVar))
  {
    agg_fpvar[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),mean,na.rm=T)
    agg_fpvar_se[[j]][[i]] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),SEM,na.rm=T)
    agg_fpvar[[j]][[i]][,"N"] = aggregate(fp[[i]][,fpVar[j]],by=list(sex=fp[[i]][,"sex"],imp=ml$where[,fpVar[j]],age_cut=fp[[i]][,"age_cut"]),length)[,"x"]
    agg_fpvar[[j]][[i]][,"sex"] = as.numeric(factor(agg_fpvar[[j]][[i]][,"sex"]),c("male","female"))-1
    agg_fpvar_se[[j]][[i]][,"N"] = 0
    agg_fpvar_se[[j]][[i]][,"sex"] = 0
  }

}
fp = do.call(rbind,fp)
C = RubinMat(C)
Cv = RubinMat(Cv)
Cvi = RubinMat(Cvi)
mu = RubinMat(mu)

agg = RubinMat(agg,lse=aggse)
vars = c(fpVar,"Nfp","Nfp_noimp")
agg[[1]] = as.data.frame(agg[[1]])
agg[[1]][,sprintf("%s_se",vars)] = agg[[2]][,vars]
agg = agg[[1]]

for (j in 1:length(fpVar))
{
  agg_fpvar[[j]] = RubinMat(agg_fpvar[[j]],agg_fpvar_se[[j]])
  agg_fpvar[[j]][[1]] = as.data.frame(agg_fpvar[[j]][[1]])
  agg_fpvar[[j]][[1]][,sprintf("%s_se","x")] = agg_fpvar[[j]][[2]][,"x"]
  agg_fpvar[[j]] = agg_fpvar[[j]][[1]]
}


library(mgcv)
gm =  list()
testage = seq(min(long[,"age"],na.rm=T)-1,max(long[,"age"],na.rm=T)+1,lenght=101)
for (j in 1:length(agg_fpvar))
{
  fitdata = subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0 & x > 0 & x_se > 0)
  #mexp = lm(log(x)~age_cut,fitdata,weights=1/fitdata[,"x_se"]^2)
  #g = gam(x~s(age_cut,3),data=fitdata,weights=1/fitdata[,"x_se"]^2)
  
  Nimp = sum(ml$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gm[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==0),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gm[[j]] = gm[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth(data=subset(agg_fpvar[[j]], N > 1 & sex==0 & imp == 0),mapping=aes(x=age_cut,y=x),inherit.aes=F,se=F,lty=2,method="gam",formula=y~s(x,4))+
  #geom_line(data=data.frame(age=testage,y=exp(predict(mexp,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  #geom_line(data=data.frame(age=testage,y=exp(predict(g,data.frame(age_cut=testage)))),aes(x=age,y=y),inherit.aes=F,lty=2,colour="black")+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1,fullrange=T)+
  labs(x="Age",title=sprintf("%s (males)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}

gm[[length(gm)+1]] = ggplot(data=subset(agg,sex==0 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  ##scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gm,nrow=2,ncol=3,top="males")

gf =  list()
for (j in 1:length(agg_fpvar))
{
    fitdata = subset(agg_fpvar[[j]], N > 1 & sex==1 & imp == 0 & x > 0 & x_se > 0)
  Nimp = sum(ml$where[,fpVar[j]])
  if(Nimp > 1)
  {
    gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se,color=ordered(imp)))
  }
  else gf[[j]] = ggplot(data=subset(agg_fpvar[[j]], N > 1 & sex==1),aes(x=age_cut,y=x,ymin=x-x_se,ymax=x+x_se))
  gf[[j]] = gf[[j]] +  geom_pointrange(position=position_dodge(1))+
  #geom_smooth(method="lm",data=subset(agg_fpvar[[j]],N>1 & imp==0),aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,fullrange =T)+
  #geom_smooth()+
  geom_smooth(data=fitdata,aes(x=age_cut,y=x,weight=1/x_se^2),lty=2,inherit.aes=F,se=F,size=1)+
  labs(x="Age",title=sprintf("%s (females)",fpVar[j]))+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()
}


#females
gf[[length(gf)+1]] = ggplot(data=subset(agg,sex==1 & N > 1),aes(x=age_cut,y=Nfp,ymin=Nfp-Nfp_se,ymax=Nfp+Nfp_se,color=ordered(Nimp)))+
  geom_pointrange()+
  geom_smooth(se=F)+
  #scale_y_log10()+
  #annotation_logticks(sides="l")+
  scale_color_discrete_diverging(palette="Blue-Red")+ #Blue-Red 3
  theme_minimal()


marrangeGrob(gf,nrow=2,ncol=3,top="females")


save=T
if(save)
{   
    ggsave(sprintf("%s/results/elsa_cart_males_fp.pdf",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/elsa_cart_males_fp.png",outputDir),
           marrangeGrob(gm,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    
      ggsave(sprintf("%s/results/elsa_cart_females_fp.pdf",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
    ggsave(sprintf("%s/results/elsa_cart_females_fp.png",outputDir),
           marrangeGrob(gf,nrow=2,ncol=3,top=NULL),width=16,height=9,dpi=300)
}
```

#
# Missingness statistics
#
```{r}
plotdata = data.frame(missingness = apply(is.na(long),2,mean),names=colnames(long),
                      N = nrow(long)
                      )
#not sure what error is...
#https://www.researchgate.net/post/Can-standard-deviation-and-standard-error-be-calculated-for-a-binary-variable
#sqrt(p.est*(1-p.est)/n)
plotdata[,"se"] = sqrt(plotdata[,"missingness"]*(1-plotdata[,"missingness"])/plotdata[,"N"])
plotdata[,"names"] = factor(plotdata[,"names"],plotdata[sort.list(plotdata[,"missingness"],decreasing=T),"names"])
ggplot(plotdata,aes(x=names,y=missingness,ymin=missingness-se,ymax=missingness+se))+
  geom_pointrange()+
  labs(x="")+
  theme_minimal()+
  theme(axis.text.x=element_text(angle=90,vjust=.5))
```

Just a few variables dominate the missingness, namely weight and gait speed.

```{r}
#relation between missingness and health
plotdata = data.frame(missing = apply(is.na(long),1,sum))

plotdata[,"fi"] = apply(long[,fiVar],1,mean,na.rm=T)
plotdata[,"age"] = long[,"age"]
plotdata[,"fic"] = cut(plotdata[,"fi"],c(-Inf,.1,.21,.45,Inf)) #blodgett cuts
plotdata[,"agec"] = cut(plotdata[,"age"],c(0,60,70,80,Inf))
ggplot(plotdata,aes(x=fi,y=missing,colour=agec,shape=agec))+geom_point()+geom_smooth()
```

No obvious connection between FI, age and missingness.

```{r}
#relation between missingness and health
plotdata = data.frame(missingNFP = apply(is.na(long[,fpVar]),1,sum))

plotdata[,"fi"] = apply(long[,fiVar],1,mean,na.rm=T)
plotdata[,"age"] = long[,"age"]
plotdata[,"fic"] = cut(plotdata[,"fi"],c(-Inf,.1,.21,.45,Inf)) #blodgett cuts
plotdata[,"agec"] = cut(plotdata[,"age"],c(0,60,70,80,Inf))
ggplot(plotdata,aes(x=age,y=missingNFP,colour=fic,shape=fic))+stat_summary()+geom_smooth()
```


Looks like people missing FP variables are much more likely to have high FI and slightly more likely to be older.

#
# Check Imputation quality
#

```{r}
#important variables - binary only
library(ggplot2)
vars = c(fpVar,"fi_prev") #binary with highest missingness + other FP vars + FI (since in range 0,1)
plotdata = list()
ml = mcart
for (i in 1:ml$m)
{
  im = mice::complete(ml,i)
  subdata = list()
  for (j in 1:length(vars))
  {
    subdata[[j]] = data.frame(variable=vars[j],value=im[,vars[j]],imputed=ml$where[,vars[j]],imputation=j)
    ind = ind + 1
  }
  subdata = do.call(rbind,subdata)
  plotdata[[i]] = as.matrix(subdata[,-1])
}
plotdata = RubinMat(plotdata)[[1]]
plotdata = data.frame(variable=subdata[,1],plotdata)
plotdata[,"imputed"] = plotdata[,"imputed"] > 0.5

ggplot(plotdata,aes(x=variable,y=value,fill=imputed,colour=imputed,shape=imputed))+
  #geom_boxplot()+ #just 0 and 1
  stat_summary(position=position_dodge(.1))+
  theme_minimal()+
  theme(axis.text.x=element_text(angle=90,vjust=.5))
```

Imputed values waaaaay more likely to be high. So why is imputed FI more likely to be low?

```{r}
#important variables
library(ggplot2)
vars = c(fpVar,"fi_prev","weight","height") #continuous with highest missingness
plotdata = list()

mu = apply(long[,vars],2,mean,na.rm=T)
s = apply(long[,vars],2,sd,na.rm=T)
ml = mcart
for (i in 1:ml$m)
{
  im = mice::complete(ml,i)
  subdata = list()
  for (j in 1:length(vars))
  {
    subdata[[j]] = data.frame(variable=vars[j],value=(im[,vars[j]]-mu[j])/s[j],imputed=ml$where[,vars[j]],imputation=j)
    ind = ind + 1
  }
  subdata = do.call(rbind,subdata)
  plotdata[[i]] = as.matrix(subdata[,-1])
}
plotdata = RubinMat(plotdata)[[1]]
plotdata = data.frame(variable=subdata[,1],plotdata)
plotdata[,"imputed"] = plotdata[,"imputed"] > 0.5

ggplot(plotdata,aes(x=variable,y=value,fill=imputed,colour=imputed,shape=imputed))+
  #geom_boxplot()+ #just 0 and 1
  stat_summary(position=position_dodge(.1))+
  labs(x="",y="Value (standardized)")+
  theme_minimal()+
  theme(axis.text.x=element_text(angle=90,vjust=.5))
```
```{r}
# NFP
im = mice::complete(ml,1)
plotdata = data.frame(Nfp = apply(im[,fpVar],1,sum,na.rm=T),Nimp = apply(ml$where[,fpVar],1,sum) ,age=im[,"age"],sex=im[,"sex"])
print(quantile(plotdata[,"Nimp"]))
plotdata[,"agebin"] = cut(plotdata[,"age"],c(0,seq(40,100,by=5),Inf))
ggplot(plotdata,aes(x=agebin,y=Nfp,colour=Nimp > 1))+stat_summary(position=position_dodge(.1))+geom_smooth()
```
Imputed values are clearly more likely to be abnormal.

```{r}
# FI
im = mice::complete(ml,1)
plotdata = data.frame(fi = apply(im[,fiVar],1,mean,na.rm=T),Nimp = apply(ml$where[,fpVar],1,sum) ,age=im[,"age"],sex=im[,"sex"])
plotdata[,"agebin"] = cut(plotdata[,"age"],c(0,seq(40,100,by=5),Inf))
ggplot(plotdata,aes(x=agebin,y=fi,colour=Nimp > 1))+
  labs(x="Age",y="FI",colour="NFP Missing")+
  stat_summary(position=position_dodge(.1))+
  geom_smooth()
```

Looks like people missing FP variables are less healthy, consistent with previous plot. SEVERE MNAR!

# 
# Wide imputation
#

```{r}
stop("not sure I want to bother")
```

# convert data to wide format
```{r}
#unique IDs
unid = elsa[[1]][,"id"]
for (i in 1:length(elsa)) unid = c(unid,elsa[[i]][,"id"])
unid = unique(unid)
```

```{r}
wide = list()
for (i in 1:length(elsa)) 
{
  wide[[i]] = data.frame(matrix(NA,nrow=length(unid),ncol=length(fiVar)+length(fpVar)+length(auxVar)))
  rownames(wide[[i]]) = unid
  wide[[i]][rownames(elsa[[i]]),] = elsa[[i]][,c(fiVar,fpVar,auxVar,"fi_prev")]
  colnames(wide[[i]]) = c(fiVar,fpVar,auxVar,"fi_prev")
  colnames(wide[[i]]) = sprintf("%s_%s",colnames(wide[[i]]),names(elsa)[[i]])
}
wide = do.call(cbind,wide)
```

```{r}
#constant variables
sex = apply(wide[,sprintf("sex_%s",names(elsa))],1,mean,na.rm=T)
wide[,"sex"] = 1*(sex > .5)

#drop extra sex columns
wide = wide[,setdiff(colnames(wide),sprintf("sex_%s",names(elsa)))]
```


```{r}
ml = mice::mice(wide,m=15,method="cart") 
```

```{r}
saveRDS(m,sprintf("%s/data/elsa_fi_vs_fp_cart15.rds",outputDir))
```